---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

## Loading and preprocessing the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "./r/activity.zip",method = "curl")
unzip("./r/activity/activity.zip", exdir = "./r/activity/")
amdata<- read.csv("./r/activity/activity.csv", stringsAsFactors = FALSE)
amdata$date<- as.Date(amdata$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(data = amdata,
  aes(date, steps)) +
  stat_summary(fun.y = sum,
    geom = "bar") 
```

Calculate and report the mean and median total number of steps taken per day
```{r}
q3sum <- aggregate(steps ~ date, amdata, sum)
q3mean <- aggregate(steps ~ date, amdata, mean)
q3median <- aggregate(steps ~ date, amdata, median, na.rm=TRUE)
q3data <- merge(q3sum, q3mean, by = "date")
q3data <- merge(q3data, q3median, by = "date")
q3data <- setNames(q3data, c("Date", "Total steps", "Mean", "Median"))
q3data
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis). Time format for interval is not changed: 0 => 00:00, 500 => 05:00, 1500 => 15:00
```{r}
cleandata <- amdata[!is.na(amdata$steps),] ## exclude NA values
q4data <- aggregate(steps~interval, data=cleandata, FUN=mean) 
library(ggplot2)
qplot(x=interval, y=steps, data = q4data, geom="line")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
q4data[max(q4data$steps),]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
``` {r}
sum(is.na(amdata$steps))
```
Strategy for filling in all of the missing values in the dataset: average for each 5-minute interval across all days. Create a new dataset that is equal to the original dataset but with the missing data filled in.
``` {r}
q6data <- amdata
for(i in which(is.na(q6data[, "steps"]))) {
  q6data$steps[i] <- q4data[q6data$interval[i]==q4data$interval,"steps"]
}
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
library(ggplot2)
ggplot(data = q6data,
  aes(date, steps)) +
  stat_summary(fun.y = sum, # adds up all observations for the date
    geom = "bar") +
  geom_line()
```
``` {r}
q7sum <- aggregate(steps ~ date, q6data, sum)
q7mean <- aggregate(steps ~ date, q6data, mean)
q7median <- aggregate(steps ~ date, q6data, median)
q7data <- merge(q7sum, q7mean, by = "date")
q7data <- merge(q7data, q7median, by = "date")
q7data <- setNames(q7data, c("Date", "Total steps", "Mean", "Median"))
q7data
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
q8data<- q6data
library(lubridate)
q8data$date<- as.POSIXct(q6data$date, format="%Y-%m-%d")
q8weekends<- q8data[wday(q8data$date) %in% 6:7,]
q8weekdays<- q8data[wday(q8data$date) %in% 1:5,]
```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
library(ggplot2)
ggplot(data = q8weekdays,
  aes(interval, steps)) +
  stat_summary(fun.y = mean, 
  geom = "bar")+
  ggtitle("Weekdays")
```
```{r}
library(ggplot2)
ggplot(data = q8weekends,
  aes(interval, steps)) +
  stat_summary(fun.y = mean, 
    geom = "bar")+
  ggtitle("Weekends") 
```
